KidsAIR wood stove grades and usage

### 2020-06-24 stove_grade_usage_cleaning.Rmd
check home char across winters

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 15 Jan 2020"
output: pdf_document
---

@@

##### WMT #####
# Load data
stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_stove_grades_20200113.csv"))

wmt_stove_grades <- stove_grades_initial_wmt %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(`REPLACE(REPLACE(Notes`),
         area = "WMT") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(wmt_stove_grades, "Output/wmt_stove_grades.rds")

##### NN #####
# Load data
stove_grades_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_stove_grades_20200113.xlsx"))

nn_stove_grades <- stove_grades_initial_nn %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(IgnoreReason),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(nn_stove_grades, "Output/nn_stove_grades.rds")

##### AK #####
# Load data
stove_grades_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_stove_grades_20200113.xlsx"))

ak_stove_grades <- stove_grades_initial_ak %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(IgnoreReason),
         area = "AK") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

@@

# Join stove grade data from different areas
```{r}
wmt_stove_grades <- read_rds("Output/wmt_stove_grades.rds")
nn_stove_grades <- read_rds("Output/nn_stove_grades.rds")
ak_stove_grades <- read_rds("Output/ak_stove_grades.rds")

@@

# write_rds(kids_stove_grades, "Output/kids_stove_grades.rds")
```

# Clean up all stove grade data
```{r}
kids_stove_grades <- read_rds(paste0(file_path, "Output/kids_stove_grades.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

stove_grades_clean <- kids_stove_grades %>% 
  filter(home_id != 0) %>% 
  filter(stove_grade != "D") %>% 
  rename(stove_grade_comments = comments) %>% 
  select(-home_id) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
  arrange(area, home, winter_id)

@@

# Dataset showing home with no stove grade
stove_grades_na <- stove_grades_clean %>% 
  filter(is.na(stove_grade)) %>% 
  distinct(home_id) 
```

@@

# Load, format, and save wood stove use data
```{r}
##### WMT #####
# Load data
stove_use_initial_wmt <- read_csv("Input/WMT/wmt_stove_use_20200113.csv")

@@

# Save as RDS
# write_rds(wmt_stove_use, "Output/wmt_stove_use.rds")

##### NN #####
# Load data
stove_use_initial_nn <- read_xlsx("Input/NN/nn_stove_use_20200113.xlsx")

@@

# Save as RDS
# write_rds(nn_stove_use, "Output/nn_stove_use.rds")

##### AK #####
# Load data
stove_use_initial_ak <- read_xlsx("Input/AK/ak_stove_use_20200113.xlsx")

@@

# Save as RDS
# write_rds(ak_stove_use, "Output/ak_stove_use.rds")
```

# Join stove use data from different areas
```{r}
wmt_stove_use <- read_rds("Output/wmt_stove_use.rds")
nn_stove_use <- read_rds("Output/nn_stove_use.rds")
ak_stove_use <- read_rds("Output/ak_stove_use.rds")

@@

# write_rds(kids_stove_use, "Output/kids_stove_use.rds")
```

# Clean up all stove use data
```{r}
kids_stove_use <- read_rds(paste0(file_path, "Output/kids_stove_use.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

stove_use_clean <- kids_stove_use %>% 
  replace_with_na(replace = list(wood_cords = -999)) %>%
  replace_with_na(replace = list(wood_cords = 0)) %>%
  mutate(wood_collect = factor(wood_collect,
                               levels = c("Less than 1 week",
                                          "Between 1 week and 1 month",
                                          "Between 1 month and 3 months",
                                          "Between 3 months and 6 months",
                                          "Between 6 months and 1 year",
                                          "More than 1 year"),
                               labels = c("< 1 week",
                                          "1 week to 1 mo",
                                          "1 mo to 3 mo",
                                          "3 mo to 6 mo",
                                          "6 mo to 1 year",
                                          "> 1 yr"))) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))

>>>>

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 23 June 2020"
output: pdf_document
---

@@

##### WMT #####
# Load data
stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/stove.csv"), 
                                  na = c("NULL", ""))

wmt_stove_grades <- stove_grades_initial_wmt %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "WMT") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
write_rds(wmt_stove_grades, paste0(file_path, "Output/wmt_stove_grades.rds"))

##### NN #####
# Load data
stove_grades_initial_nn <- read_csv(paste0(file_path, "Input/NN/stove.csv"), 
                                  na = c("NULL", ""))

nn_stove_grades <- stove_grades_initial_nn %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
write_rds(nn_stove_grades, paste0(file_path, "Output/nn_stove_grades.rds"))

##### AK #####
# Load data
stove_grades_initial_ak <- read_csv(paste0(file_path, "Input/AK/stove.csv"), 
                                  na = c("NULL", ""))

ak_stove_grades <- stove_grades_initial_ak %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "AK") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

@@

# Join stove grade data from different areas
```{r}
wmt_stove_grades <- read_rds(paste0(file_path, "Output/wmt_stove_grades.rds"))
nn_stove_grades <- read_rds(paste0(file_path, "Output/nn_stove_grades.rds"))
ak_stove_grades <- read_rds(paste0(file_path, "Output/ak_stove_grades.rds"))

@@

write_rds(kids_stove_grades, paste0(file_path, "Output/kids_stove_grades.rds"))
```

# Clean up all stove grade data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_stove_grades <- read_rds(paste0(file_path, "Output/kids_stove_grades.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

stove_grades_clean <- kids_stove_grades %>% 
  mutate(area = as.factor(area)) %>% 
  filter(home_id != 0) %>% 
  # Mark gave a couple of "D" grades when there was not enough info to score the stove
  # removing those "D" grades from final dataset
  filter(stove_grade != "D") %>% 
  select(-home_id) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
  arrange(area, home, winter_id)

summary(stove_grades_clean)

# Some homes/stoves were graded during both winters, typically when new photos were added
# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
# We can't make the assumption that the stove was exactly the same for both winters,
# so it will be best to use the grade corresponding to each winter for analysis.

@@

# Dataset showing homes with no stove grade
stove_grades_na <- stove_grades_clean %>% 
  filter(is.na(stove_grade)) %>% 
  distinct(home_id) 
```

@@

# Load, format, and save wood stove use data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
stove_use_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

@@

# Save as RDS
write_rds(wmt_stove_use, paste0(file_path, "Output/wmt_stove_use.rds"))

##### NN #####
# Load data
stove_use_initial_nn <- read_csv(paste0(file_path, "Input/NN/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

@@

# Save as RDS
write_rds(nn_stove_use, paste0(file_path, "Output/nn_stove_use.rds"))

##### AK #####
# Load data
stove_use_initial_ak <- read_csv(paste0(file_path, "Input/AK/woodstoveusage.csv"), 
                                  na = c("NULL", ""))

@@

# Save as RDS
write_rds(ak_stove_use, paste0(file_path, "Output/ak_stove_use.rds"))
```

# Join stove use data from different areas
```{r}
wmt_stove_use <- read_rds(paste0(file_path, "Output/wmt_stove_use.rds"))
nn_stove_use <- read_rds(paste0(file_path, "Output/nn_stove_use.rds"))
ak_stove_use <- read_rds(paste0(file_path, "Output/ak_stove_use.rds"))

@@

write_rds(kids_stove_use, paste0(file_path, "Output/kids_stove_use.rds"))
```

# Clean up all stove use data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_stove_use <- read_rds(paste0(file_path, "Output/kids_stove_use.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

stove_use_clean <- kids_stove_use %>% 
  mutate(area = as.factor(area)) %>% 
  replace_with_na(replace = list(wood_cords = -999)) %>%
  replace_with_na(replace = list(wood_cords = 0)) %>%
  mutate(wood_collect = factor(wood_collect,
                               levels = c("Less than 1 week",
                                          "Between 1 week and 1 month",
                                          "Between 1 month and 3 months",
                                          "Between 3 months and 6 months",
                                          "Between 6 months and 1 year",
                                          "More than 1 year"),
                               labels = c("< 1 week",
                                          "1 week to 1 mo",
                                          "1 mo to 3 mo",
                                          "3 mo to 6 mo",
                                          "6 mo to 1 year",
                                          "> 1 yr"))) %>% 
  right_join(kids_linked_ids, by = c("area", "home_winter_id"))

summary(stove_use_clean)
